---
title: "Combined analysis of Near/Mid tasks from Study 1 and Study 2"
author: "George Kinnear"
date: "08/09/2021"
output: 
  html_document:
    code_folding: hide
    df_print: paged
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)

#Plots
library(ggplot2)
library(patchwork) # for combining plots
library(ggtext) # for markdown in plots
library(latex2exp) # for formatting latex in plot captions

# For Bayesian stats:
library(bayestestR)
#library(insight)
#library(see)
library(rstanarm)
library(tidybayes)
#library(BayesFactor)
ci_width = 0.95  # Use 95% HDIs
set.seed(20210518)

# Tables
library(knitr)
library(kableExtra)
basic_kable = function(df, ...) {
  df %>% 
    kable(...) %>%
    kable_styling(bootstrap_options = "striped", full_width = F)
}

# Plot styling
heathers = c("#00b3c3","#e51c24","#016637","#fcb123")
heathers = c("#87005a", "#c66005","#016637","#fcb123")
groupcolours = c("C-CG" = heathers[1],
                 "C-GC" = "#6f004a", # darker heathers1
                 "G-CG" = heathers[2],
                 "G-GC" = "#a34e03" # darker heathers2
                 )
four_groups = c("CCG", "CGC", "GCG", "GGC")
theme_set(theme_minimal())
```

The Near and Mid tasks were unchanged from Study 1 to Study 2, so we combine the student responses from both studies to analyse possible interactions between the tasks, i.e.

> is there any evidence of task order influencing students' performance?

```{r read-data, message=FALSE, warning=FALSE}
question_marks = read_csv("data-clean/study1/FAC_LGE-tasks.csv") %>%
  select(question_name, marks_available)

participants <- bind_rows(
  read_csv("data-clean/study1/participants.csv") %>% 
    mutate(
      group = str_replace(group, "LGE Group ", ""),
      group = str_glue("study1_{str_sub(group, 1, 1)}-{str_sub(group, 2, 3)}"),
      username = paste0("study1_", username)
    ),
  read_csv("data-clean/study2/participants.csv") %>% 
    mutate(
      group = paste0("study2_", str_replace(group, "Group ", "")),
      username = paste0("study1_", username)
    )
)

attempt_data_all = read_rds("data-clean/FAC_LGE-attempt_data.rds")

task_order_data <- bind_rows(
    "study1" = read_rds("data-clean/FAC_LGE-attempt_data.rds"),
    "study2" = read_rds("data-clean/FAC_LGE2-attempt_data.rds"),
    .id = "study"
  ) %>% 
  filter(
    str_detect(question_name, "Near|Mid"),
    attempt_number == 1
  ) %>% 
  left_join(question_marks, by = "question_name") %>% 
  mutate(
    # default to 1 mark per question if not specified
    marks_available = replace_na(marks_available, 1),
    marks_attained = fraction * marks_available,
    # check the time of the attempt to see if it was before the end of the study
    # (since on 2019-11-27, all versions of the materials were made available to all students)
    attempt_time = lubridate::as_datetime(timecreated),
    during_study = case_when(
      study == "study1" ~ attempt_time < lubridate::ymd("2019-11-27"),
      study == "study2" ~ attempt_time < lubridate::ymd("2020-11-27"),
    ),
    group = str_remove_all(group, "LGE Group |Group ")
  ) %>% 
  filter(during_study, !is.na(marks_attained)) %>% 
  mutate(q_type = if_else(str_detect(question_name, "C"), "C", "G")) %>% 
  select(study, group, username, question_name, q_type, questionsummary, responsesummary, marks_attained) %>% 
  mutate(task_order = if_else(str_detect(group, "CG$"), "Classify first", "Generate first")) 

task_order_data %>% 
  select(study, task_order, username) %>% 
  distinct() %>% 
  janitor::tabyl(study, task_order) %>% 
  janitor::adorn_totals(where = "both") %>% 
  basic_kable()
```


# Item details

For each item, we have full details of the student response and how it was graded by STACK. Here is an example for each of the tasks:

```{r message=FALSE, warning=FALSE}
responses_raw <- task_order_data

responses_raw %>% group_by(question_name) %>% slice_head() %>% basic_kable()
```

We process this raw data to extract the answers to each part, and the score.

```{r}
responses <- responses_raw %>% 
  # pick out the various answers - str_match_all returns a matrix, which we have to wrangle a bit
  mutate(ans_bits = str_match_all(responsesummary, "(ans[^:]*): ([^;]*)(?=\\[score\\];)")) %>% 
  #mutate(ans_bits = str_match_all(responsesummary, "(ans[:digit:]): ([^;]*)(?=\\[score\\];)")) %>% 
  mutate(ans_bits = map(ans_bits, ~ tibble(part = .x[,2], response = .x[,3]))) %>% 
  unnest(ans_bits) %>% 
  mutate(response = str_remove_all(response, '"')) %>% 
  # construct a simplified name for each question part
  mutate(q_part = paste(question_name, part, sep = "_")) %>% 
  mutate(
    relevant_prt = case_when(
      # pick out the PRT trace that is relevant to each part, i.e. that grades that answer
      # for these questions, ansN corresponds to prtN in all cases
      TRUE ~ str_match_all(responsesummary, str_glue("prt{parse_number(part)}: [^;]*(?=; |$)")) %>% as.character()
    ),
    fully_correct = case_when(
      # the classification tasks all have very simple PRTs with a single node deciding the mark
      str_detect(question_name, "C") ~ str_detect(relevant_prt, "-1-T"),
      # MidG also has quite simple PRTs
      q_part %in% c("MidG_ans1", "MidG_ans2", "MidG_ans3") ~ str_detect(relevant_prt, "-2-T"),
      q_part == "MidG_ans4" ~ str_detect(relevant_prt, "-1-T"),
      # NearG should be increasing (-1-T) and bounded (-4-T)
      question_name == "NearG" ~ str_detect(relevant_prt, "-1-T") & str_detect(relevant_prt, "-4-T"),
      TRUE ~ NA
    )
  )
```

```{r}
# responses %>% 
#   group_by(question_name, part) %>% 
#   summarise(
#     n = n(),
#     num_correct = sum(fully_correct),
#     num_incorrect = sum(!fully_correct),
#     prop_correct = num_correct / n,
#     .groups = "drop"
#   ) %>% 
#   pivot_longer(cols = starts_with("num_"), names_to = "type", values_to = "count", names_prefix = "num_") %>% 
#   mutate(signed_count = if_else(type == "correct", count, -count)) %>% 
#   mutate(question_name = fct_relevel(question_name, "NearC", "NearG", "MidC1", "MidC2", "MidG", "FarC1", "FarC2")) %>% 
#   mutate(section = str_extract(question_name, "Near|Mid|Far") %>% fct_relevel("Near", "Mid", "Far")) %>% 
#   mutate(q_part = fct_rev(paste0(str_extract(question_name, "G|C[:digit:]?"), letters[parse_number(part)]))) %>% 
#   mutate(part = fct_rev(letters[parse_number(part)])) %>% 
#   ggplot(aes(x = q_part, y = signed_count, fill = type)) +
#   geom_bar(position = "stack", stat = "identity") +
#   geom_hline(yintercept = 0) +
#   coord_flip() +
#   scale_y_continuous("Number of responses", labels = abs, breaks = seq(-50, 100, 50)) +
#   scale_fill_viridis_d("Answer", option = "plasma", end = 0.8, direction = -1, guide = guide_legend(reverse = TRUE)) +
#   facet_grid(rows = vars(section), scales = "free", space = "free", switch = "both") +
#   theme(
#     strip.placement = "outside",
#     strip.text.y.left = element_text(angle = 0),
#     legend.position = "top"
#   ) +
#   labs(x = "")

responses %>% 
  group_by(question_name, part) %>% 
  summarise(
    n = n(),
    num_correct = sum(fully_correct),
    num_incorrect = sum(!fully_correct),
    prop_correct = num_correct / n,
    .groups = "drop"
  ) %>% 
  pivot_longer(cols = starts_with("num_"), names_to = "type", values_to = "count", names_prefix = "num_") %>% 
  mutate(signed_count = if_else(type == "correct", count, -count)) %>% 
  mutate(question_name = fct_relevel(question_name, "NearC", "NearG", "MidC1", "MidC2", "MidG", "FarC1", "FarC2")) %>% 
  mutate(part = fct_rev(letters[parse_number(part)])) %>% 
  ggplot(aes(x = part, y = signed_count, fill = type)) +
  geom_bar(position = "stack", stat = "identity") +
  geom_hline(yintercept = 0) +
  coord_flip() +
  scale_y_continuous("Number of responses", labels = abs, breaks = seq(-50, 100, 50)) +
  scale_fill_viridis_d("Answer", option = "plasma", end = 0.8, direction = -1, guide = guide_legend(reverse = TRUE)) +
  facet_grid(rows = vars(question_name), scales = "free", space = "free", switch = "both") +
  theme(
    strip.placement = "outside",
    strip.text.y.left = element_text(angle = 0),
    legend.position = "top"
  ) +
  labs(x = "")

responses %>% 
  mutate(study_group = if_else(study == "study1", str_glue("Study 1 ({str_sub(group, start = 1, end = 1)}-)"), "Study 2")) %>% 
  group_by(study_group, question_name, part) %>% 
  summarise(
    n = n(),
    num_correct = sum(fully_correct),
    num_incorrect = sum(!fully_correct),
    prop_correct = num_correct / n,
    .groups = "drop"
  ) %>% 
  pivot_longer(cols = starts_with("num_"), names_to = "type", values_to = "count", names_prefix = "num_") %>% 
  mutate(signed_count = if_else(type == "correct", count, -count)) %>% 
  mutate(question_name = fct_relevel(question_name, "NearC", "NearG", "MidC1", "MidC2", "MidG", "FarC1", "FarC2")) %>% 
  mutate(part = fct_rev(letters[parse_number(part)])) %>% 
  ggplot(aes(x = part, y = count, fill = type)) +
    geom_bar(position = "stack", stat = "identity") +
    geom_hline(yintercept = 0) +
    coord_flip() +
    #scale_y_continuous("Number of responses", labels = abs, breaks = seq(-50, 100, 50)) +
    scale_y_continuous("Number of responses", breaks = seq(0, 100, 20)) +
    scale_fill_viridis_d("Answer", option = "plasma", end = 0.8, direction = -1, guide = guide_legend(reverse = TRUE)) +
    facet_grid(cols = vars(study_group), rows = vars(question_name), scales = "free", space = "free", switch = "y") +
    theme(
      strip.placement = "outside",
      strip.text.y.left = element_text(angle = 0),
      legend.position = "top"
    ) +
    labs(x = "")
ggsave("figs/FIG_study2_subsequent-task-scores.eps", device = cairo_ps, width = 15, height = 9, units = "cm")
```

```{r}
task_order_scores <- responses %>% 
  select(study, task_order, username, question_name, q_type, marks_attained) %>% 
  distinct() %>% 
  group_by(study, task_order, username, q_type) %>% 
  tally(marks_attained, name = "marks_attained") %>%
  pivot_wider(names_from = "q_type", values_from = "marks_attained", values_fill = 0)
```

There are several students who did not complete any of the relevant tasks, so they are not relevant to the question:

```{r}
participants %>% 
  separate(username, into = c("study", "username"), sep = "_") %>% 
  anti_join(task_order_data %>% select(study, username) %>% distinct(), by = c("study", "username")) %>% 
  mutate(task_order = if_else(str_detect(group, "CG$"), "Classify first", "Generate first")) %>% 
  janitor::tabyl(task_order) %>% 
  basic_kable(caption = "Students who did not answer any of the Near or Mid tasks")
```

So we stick with just the `r nrow(task_order_scores)` students who completed at least one of the Near/Mid tasks.

```{r}
task_order_scores %>% write_csv("data-clean/studyX_task_order_scores.csv")

task_order_model_C <- stan_glm(
    formula = C ~ 0 + task_order,
    data = task_order_scores,
    seed = 123,
    refresh = 0 # to prevent messages about the MCMC iterations being printed
  )
means_C <- modelbased::estimate_means(task_order_model_C, ci = ci_width, centrality = "median", ci_method = "hdi")
contrasts_C <- modelbased::estimate_contrasts(task_order_model_C, ci = ci_width, centrality = "median", ci_method = "hdi")

task_order_model_G <- stan_glm(
    formula = G ~ 0 + task_order,
    data = task_order_scores,
    seed = 123,
    refresh = 0 # to prevent messages about the MCMC iterations being printed
  )
means_G <- modelbased::estimate_means(task_order_model_G, ci = ci_width, centrality = "median", ci_method = "hdi")
contrasts_G <- modelbased::estimate_contrasts(task_order_model_G, ci = ci_width, centrality = "median", ci_method = "hdi")

means = bind_rows("C" = means_C, "G" = means_G, .id = "q_type")
contrasts = bind_rows("C" = contrasts_C, "G" = contrasts_G, .id = "q_type")

task_order_scores %>% 
  pivot_longer(cols = c("G", "C"), names_to = "q_type", values_to = "marks_attained") %>% 
  ggplot(aes(x = task_order, y = marks_attained)) +
  ggbeeswarm::geom_beeswarm(alpha = 0.4, stroke = 0) +
  geom_pointrange(data = means, aes(x = task_order, y = Mean, ymin = CI_low, ymax = CI_high),
                  position = position_nudge(x = 0.35),
                  size = 0.4) +
  facet_grid(cols = vars(q_type), labeller = labeller(q_type = c("C" = "Classification score", "G" = "Generation score"))) +
  labs(y = "", x = "") +
  theme(panel.spacing.x = unit(2, "lines"))

ggsave("figs/FIG_studyX_task_order.eps", device = cairo_ps, width = 12, height = 7, units = "cm")
```

```{r}
task_order_means_table <- means %>% 
  left_join(
    task_order_scores %>% 
      pivot_longer(cols = c("G", "C"), names_to = "q_type", values_to = "marks_attained") %>% 
      group_by(q_type, task_order) %>% 
      summarise(
        N = n(),
        sd = sd(marks_attained),
        .groups = "drop"
      ),
    by = c("q_type", "task_order")
  ) %>%
  relocate(N, .before = Mean) %>% 
  mutate(score = if_else(q_type == "C", "Classification", "Generation"), .before = 1, .keep = "unused") %>% 
  mutate(estimate_hdi = str_glue("[{round(CI_low, 3)}, {round(CI_high, 3)}]"), .keep = "unused")

task_order_means_table %>% 
  basic_kable(digits = 3) %>% 
  collapse_rows(columns = 1, valign = "top")
```

```{r, comment=""}
task_order_means_table %>% 
  basic_kable(digits = 3, format = "latex", booktabs = T) %>% 
  cat()
```


```{r}
# inspired by https://stackoverflow.com/questions/61977439/table-below-x-axis-in-ggplot
p1 <- task_order_scores %>% 
  pivot_longer(cols = c("G", "C"), names_to = "q_type", values_to = "marks_attained") %>% 
  ggplot(aes(x = task_order, y = marks_attained)) +
  ggbeeswarm::geom_beeswarm(alpha = 0.4, stroke = 0) +
  geom_pointrange(data = means, aes(x = task_order, y = Mean, ymin = CI_low, ymax = CI_high),
                  position = position_nudge(x = 0.35),
                  size = 0.4) +
  facet_grid(cols = vars(q_type), labeller = labeller(q_type = c("C" = "Classification score", "G" = "Generation score"))) +
  labs(y = "", x = "") +
  theme(panel.spacing.x = unit(2, "lines"), strip.text = element_text(size = 11))

p2 <- task_order_means_table %>% 
  mutate(
    n = str_glue("{N}"),
    mean = str_glue("{round(Mean, digits = 3)}"),
    sd = str_glue("{round(sd, digits = 3)}")
  ) %>% 
  pivot_longer(c(n, mean, sd, estimate_hdi), names_to = "layer", values_to = "label") %>% 
  mutate(layer = fct_relevel(layer, c("n", "sd", "mean", "estimate_hdi")) %>% fct_rev()) %>% 
  #filter(layer != "n") %>% 
  ggplot(aes(x = task_order, y = layer, label = label)) +
  geom_text(size = 10 * 5/14) +
  labs(y = "", x = NULL) +
  scale_y_discrete(labels = c("n" = "N", "mean" = "Mean", "estimate_hdi" = "\n95% HDI\nfor mean")) +
  theme_minimal() +
  facet_grid(cols = vars(score), labeller = labeller(q_type = c("C" = "Classification score", "G" = "Generation score"))) +
  theme(axis.line = element_blank(), axis.ticks = element_blank(), axis.text.x = element_blank(),
        panel.grid = element_blank(), strip.text = element_blank(),
        panel.spacing.x = unit(2, "lines"))
  
p1 / p2 +  plot_layout(heights = c(8, 3))
ggsave("figs/FIG_studyX_task_order_with_table.eps", device = cairo_ps, width = 15, height = 10, units = "cm")
```

```{r}
contrasts %>% 
  basic_kable(digits = 3)
```

# Interactions between tasks?

## NearG

> Give an example of a sequence which is increasing and bounded above.

There are a couple of hypotheses about students' answers to this task, if they are being influenced by previous tasks:

1. students from study 1 G- groups will be more likely to use examples based on `1/2^n` than students from other groups, since they will have seen that as a worked solution to G1.

2. excluding the study1 G- groups, students in classify-first groups will be more likely to use `1/n`-based examples than those from generate-first groups, since they will have just seen `1/(an+b)` in NearC.

First we classify the type of response given by each student, blind to which group they are in:

```{r}
nearG_responses <- responses %>% 
  select(-questionsummary) %>% 
  filter(question_name == "NearG") %>% 
  mutate(
    response_type = case_when(
      str_detect(response, "sqrt|u|log|sin|ln") ~ "other",
      # pick off some odd cases first
      str_trim(response) %in% c("(1/2*n)^(-1)") ~ "other",
      # identify use of example from G1
      str_detect(response, "/2\\^n|\\(1/2\\)\\^n") ~ "1/2^n",
      # pick off other uses of ^n
      str_detect(response, "\\^[-]?n|\\*\\*n") ~ "^n",
      # identify examples similar to NearC
      str_detect(response, "1/n(?!\\^)") ~ "1/n", # look ahead to rule out 1/n^2 etc
      str_detect(response, "[:digit:]/[:digit:]*[\\*]?n(?!\\^)") ~ "a/n",
      str_detect(response, "[:digit:]/\\([:digit:]*[\\*]?n\\)(?!\\^)") ~ "a/n",
      str_detect(response, "n/\\(n\\+1\\)|\\(n-1\\)/n") ~ "(expr in n)/(an+b)",
      str_detect(response, ".*n.*/\\([:digit:]*[\\*]*n[\\+-]*[:digit:]*\\)") ~ "(expr in n)/(an+b)",
      str_detect(response, "/\\([:digit:]*[\\*]*n[\\+-]*[:digit:]*\\)") ~ "1/(an+b)",
      str_detect(response, "/-n") ~ "1/n",
      # other stuff
      str_detect(response, "[:digit:]*[\\*]?n[:space:]*[\\+-][:space:]*[:digit:]*") ~ "an+b",
      str_detect(response, "[:digit:]+[\\*]?n|\\([-]?[:digit:]*\\)[\\*]?n") ~ "an+b",
      str_detect(response, "n\\^[:digit:]") ~ "n^2, n^3",
      str_trim(response) %in% c("n", "1+n", "n/2") ~ "an+b",
      str_detect(str_trim(response), "^[:digit:]+$") ~ "integer",
      TRUE ~ "other"
    ),
    response_type_overall = case_when(
      response_type %in% c("1/2^n") ~ "1/2^n",
      response_type %in% c("1/n", "a/n", "(expr in n)/(an+b)", "1/(an+b)") ~ "1/n",
      TRUE ~ "other"
    )
  )

nearG_responses %>% 
  group_by(response_type_overall, response_type, response) %>% 
  tally() %>% 
  arrange(response_type_overall, response_type, -n, response) %>% 
  basic_kable() %>% 
  collapse_rows(columns = c(1:2), valign = "top")

nearG_responses %>% 
  write_csv("data-clean/studyX_NearG_responses.csv")
```

```{r}
nearG_responses %>% 
  group_by(response_type_overall, response_type) %>% 
  tally() %>% 
  arrange(response_type_overall, -n) %>% 
  basic_kable() %>% 
  collapse_rows(columns = c(1), valign = "top")
```

Now to check the hypotheses:

> 1. students from study 1 G- groups will be more likely to use examples based on `1/2^n` than students from other groups, since they will have seen that as a worked solution to G1.

```{r}
nearG_H1 <- nearG_responses %>% 
  mutate(big_group = if_else(study == "study1" & str_detect(group, "^G"), "Study 1 G-", "Other")) %>% 
  select(big_group, response_type_overall) %>% 
  janitor::tabyl(big_group, response_type_overall) %>% 
  janitor::adorn_totals(where = "col")

nearG_H1 %>% 
  basic_kable()

nearG_H1 %>% 
  pivot_longer(cols = !c(big_group, Total), names_to = "example_type", values_to = "frequency") %>% 
  ggplot(aes(x = big_group, y = frequency, fill = example_type, width = Total/100)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_viridis_d("Example type", option = "plasma", end = 0.8, direction = -1) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Group", y = "Frequency")
```

This suggests that it is indeed the case that the Study 1 G- groups are more likely to use `1/2^n`-based examples.

> 2. excluding the study1 G- groups, students in classify-first groups will be more likely to use `1/n`-based examples than those from generate-first groups, since they will have just seen `1/(an+b)` in NearC.

```{r}
nearG_H2 <- nearG_responses %>% 
  filter(!(study == "study1" & str_detect(group, "^G"))) %>% 
  mutate(big_group = if_else(str_detect(group, "G$"), "Classify first", "Generate first")) %>% 
  select(big_group, response_type_overall) %>% 
  janitor::tabyl(big_group, response_type_overall) %>% 
  janitor::adorn_totals(where = "col")

nearG_H2 %>% 
  pivot_longer(cols = !c(big_group, Total), names_to = "example_type", values_to = "frequency") %>% 
  ggplot(aes(x = big_group, y = frequency, fill = example_type, width = Total/100)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_viridis_d("Example type", option = "plasma", end = 0.8, direction = -1) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Group", y = "Frequency")

nearG_H2 %>% 
  basic_kable()
```

While the proportion is definitely higher, it's not significantly so:

```{r}
prop.test(nearG_H2$`1/n`, nearG_H2$`Total`, alternative = "greater")
```

One table and plot for both together:

```{r}
nearG_Hs <- nearG_responses %>% 
  mutate(big_group = case_when(
    study == "study1" & str_detect(group, "^G") ~ "Study 1 G-",
    str_detect(group, "G$") ~ "Classify first",
    TRUE ~ "Generate first"
  )) %>% 
  mutate(big_group = fct_relevel(big_group, "Study 1 G-")) %>% # move that group first
  select(big_group, response_type_overall) %>% 
  janitor::tabyl(big_group, response_type_overall) %>% 
  janitor::adorn_totals(where = "col") 

nearG_Hs %>% 
  pivot_longer(cols = !c(big_group, Total), names_to = "example_type", values_to = "frequency") %>% 
  ggplot(aes(x = big_group, y = frequency, fill = example_type, width = Total/100)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_viridis_d("Example type", option = "plasma", end = 0.8, direction = -1) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Group", y = "Frequency")

nearG_Hs %>% 
  janitor::adorn_percentages(,,!c(Total, big_group)) %>%
  janitor::adorn_pct_formatting(digits = 0,,,!c(Total, big_group)) %>%
  janitor::adorn_ns(position = "front",,,!c(Total, big_group)) %>% 
  rename(Group = big_group) %>% 
  relocate(N = Total, .after = Group) %>% 
  basic_kable() %>%
  add_header_above(c(" " = 2, "Example type" = 3))
```
```{r, comment=""}
nearG_Hs %>% 
  janitor::adorn_percentages(,,!c(Total, big_group)) %>%
  janitor::adorn_pct_formatting(digits = 0,,,!c(Total, big_group)) %>%
  janitor::adorn_ns(position = "front",,,!c(Total, big_group)) %>% 
  rename(Group = big_group) %>% 
  relocate(N = Total, .after = Group) %>% 
  basic_kable(format = "latex", booktabs = T) %>%
  add_header_above(c(" " = 2, "Example type" = 3)) %>% 
  cat()
```

## MidG

> In each case below, give an example of an arithmetic sequence with the stated property, by entering an expression for the general term. If it is not possible, enter `none`.
> (a) Increasing
> (b) Decreasing
> (c) Bounded above
> (d) Decreasing and bounded below

First we classify the type of response given by each student, blind to which group they are in:

```{r}
midG_responses <- responses %>% 
  select(-questionsummary) %>% 
  filter(question_name == "MidG") %>% 
  mutate(
    response_type = case_when(
      str_trim(response) %in% c("none") ~ "none",
      str_detect(response, "sqrt|u|log|sin|ln|abs") ~ "other",
      # pick off some odd cases first
      str_trim(response) %in% c("(1/2*n)^(-1)") ~ "other",
      # identify use of example from G1
      str_detect(response, "/2\\^n|\\(1/2\\)\\^n") ~ "1/2^n",
      str_detect(response, "2\\^\\(n-1\\)") ~ "1/2^n",
      str_detect(response, "1/\\(2\\^n\\)") ~ "1/2^n",
      # pick off other uses of ^n
      str_detect(response, "\\^[-]?n|\\*\\*n") ~ "^n",
      str_detect(response, "n\\^[:digit:]") ~ "n^2, n^3",
      # identify examples similar to NearC
      str_detect(response, "1/n(?!\\^)") ~ "1/n", # look ahead to rule out 1/n^2 etc
      str_detect(response, "[:digit:]/[:digit:]*[\\*]?n(?!\\^)") ~ "a/n",
      str_detect(response, "[:digit:]/\\([:digit:]*[\\*]?n\\)(?!\\^)") ~ "a/n",
      str_detect(response, "n/\\(n\\+1\\)|\\(n-1\\)/n") ~ "(expr in n)/(an+b)",
      str_detect(response, ".*n.*/\\([:digit:]*[\\*]*n[\\+-]*[:digit:]*\\)") ~ "(expr in n)/(an+b)",
      str_detect(response, "/\\([:digit:]*[\\*]*n[\\+-]*[:digit:]*\\)") ~ "1/(an+b)",
      str_detect(response, "/-n") ~ "1/n",
      # other stuff
      str_detect(response, "[:digit:]*[\\*]?n[:space:]*[\\+-][:space:]*[:digit:]*") ~ "an+b",
      str_detect(response, "[-]?[:digit:]*[:space:]*[\\+-][:space:]*[:digit:]*[\\*]?n") ~ "an+b",
      str_detect(response, "[:digit:]+[\\*]?n|\\([-]?[:digit:]*\\)[\\*]?n") ~ "an+b",
      str_trim(response) %in% c("n", "1+n", "n/2", "2-(n)", "ln(2)-ln(2)*(n-1)", "n/2+1", "n*d") ~ "an+b",
      str_detect(str_trim(response), "^[:digit:]+$") ~ "integer",
      TRUE ~ "other"
    ),
    response_type_overall = case_when(
      response_type %in% c("1/2^n") ~ "1/2^n",
      response_type %in% c("1/n", "a/n", "(expr in n)/(an+b)", "1/(an+b)") ~ "1/n",
      response_type %in% c("an+b") ~ "an+b",
      response_type %in% c("none") ~ "none",
      TRUE ~ "other"
    )
  )

midG_responses %>% 
  group_by(response_type_overall, response_type, response) %>% 
  tally() %>% 
  arrange(response_type_overall, response_type, -n, response) %>% 
  basic_kable() %>% 
  collapse_rows(columns = c(1:2), valign = "top")

midG_responses %>% 
  write_csv("data-clean/studyX_MidG_responses.csv")
```


```{r}
nearG_Hs <- midG_responses %>% 
  mutate(big_group = case_when(
    study == "study1" & str_detect(group, "^G") ~ "Study 1 G-",
    str_detect(group, "G$") ~ "Classify first",
    TRUE ~ "Generate first"
  )) %>% 
  select(big_group, response_type_overall) %>% 
  janitor::tabyl(big_group, response_type_overall) %>% 
  janitor::adorn_totals(where = "col")

nearG_Hs %>% 
  basic_kable()

midG_responses %>% 
  select(q_part, response_type_overall) %>% 
  janitor::tabyl(q_part, response_type_overall) %>% 
  janitor::adorn_totals(where = "col") %>% 
  pivot_longer(cols = !c(q_part, Total), names_to = "example_type", values_to = "frequency") %>% 
  ggplot(aes(x = q_part, y = frequency, fill = example_type)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_viridis_d("Example type", option = "plasma", end = 0.8, direction = -1) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Group", y = "Frequency")
```

How does MidC1 influence answers? It was randomised, so different students will have had examples of different types to classify:

```{r}
MidC1_examples <- responses %>% 
  filter(question_name == "MidC1") %>% 
  select(study, group, username, questionsummary) %>% 
  distinct() %>% 
  separate(questionsummary, into = c("MidC1_example", "ans"), sep = " is ") %>% 
  mutate(MidC1_example = str_remove_all(MidC1_example, "\\\\|\\(|\\)|,|\\{|\\}"))

head(MidC1_examples, n = 10) %>% 
  basic_kable()
```

The different examples used in MidC1 will be relevant to different parts of MidG:

| MidC1 example      | Gives answer to MidG |
| ------------------ | -------------------- |
| Decreasing and Bounded above | b, c |
| Increasing and Bounded below | a |

Do students make use of these examples? If so, then we'd expect to see a difference in performance on the relevant parts of MidG, based on the type of example seen in MidC1:

```{r}
midG_responses %>% 
  left_join(MidC1_examples, by = c("study", "group", "username")) %>% 
  select(q_part, response, fully_correct, MidC1_example, ans) %>% 
  mutate(ans = str_sub(ans, start = 1, end = 3)) %>% 
  group_by(q_part, ans, fully_correct) %>% 
  tally() %>% 
  filter(!is.na(ans)) %>% 
  ggplot(aes(x = ans, y = n, fill = fully_correct)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_viridis_d("Correct", option = "plasma", end = 0.8, direction = -1) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Example from MidC1", y = "Frequency") +
  facet_grid(cols = vars(q_part))
```

There appears to be no difference.

Perhaps it is restricted to just the students in the classify-first group?

```{r}
midG_responses %>% 
  left_join(MidC1_examples, by = c("study", "group", "username")) %>% 
  select(study, group, q_part, response, fully_correct, MidC1_example, ans) %>% 
  mutate(study_group = if_else(str_detect(group, "G$"), "Classify first", "Generate first")) %>% 
  mutate(ans = str_sub(ans, start = 1, end = 3)) %>% 
  group_by(study_group, q_part, ans, fully_correct) %>% 
  tally() %>% 
  filter(!is.na(ans)) %>% 
  ggplot(aes(x = ans, y = n, fill = fully_correct)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_viridis_d("Correct", option = "plasma", end = 0.8, direction = -1) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "Example from MidC1", y = "Frequency") +
  facet_grid(cols = vars(q_part), rows = vars(study_group))
```

Even here, there is no apparent difference.

Looking at the actual content of the responses:

```{r}
midGandC <- midG_responses %>% 
  left_join(MidC1_examples, by = c("study", "group", "username")) %>% 
  select(study, group, username, q_part, fully_correct, response_type, response_type_overall, response, MidC1_example, MidC1_type = ans)
```

```{r}
midGandC %>% 
  filter(q_part %in% c("MidG_ans1", "MidG_ans2", "MidG_ans3")) %>% 
  janitor::tabyl(q_part) %>% 
  janitor::adorn_totals(where = "row") %>% 
  select(-percent) %>% 
  basic_kable(caption = "Responses to relevant MidG tasks")
```


### Increasing and Bounded below

These examples from MidC1 are of exactly the right type for MidG_ans1. How often do students quote that example?

```{r}
midGandC %>% 
  filter(q_part == "MidG_ans1") %>% 
  filter(MidC1_type == "Increasing and Bounded below") %>% 
  select(study, group, username, response, MidC1_example) %>% 
  # Use the Ryacas computer algebra system to compare the two answers
  # To prepare, need to add implied multiplication to the expressions
  mutate(
    response = str_replace(response, "2u_n", "2*u_n"),
    response = str_replace(response, "\\)(?=[:digit:])", "\\)\\*"),
    response = str_replace(response, "(?<=[:digit:])\\(", "\\*\\("),
    response = str_replace(response, "(?<=[:digit:])n", "\\*n"),
    diff = paste0(response, "-(", str_replace(MidC1_example, "(?<=[:digit:])n", "\\*n"), ")")
  ) %>%  
  mutate(
    response_diff = map_chr(diff, ~ .x %>% Ryacas::y_fn("Simplify") %>% Ryacas::yac_str()),
    .keep = "unused"
  ) %>% 
  arrange(response_diff) %>% 
  basic_kable()
```

Not often at all! There are only a couple of cases of students using the identical answer (`3*n-1` and `2*n+2`), both of whom are in the study2 CG group. There are a few more cases where the answer differs only by a constant (e.g. `3n-1` becomes `3*n+1`). But overall the examples seem to be quite unrelated to MidC1.


### Decreasing and Bounded above

These examples from MidC1 are of exactly the right type for MidG_ans2 and MidG_ans3. How often do students quote that example?

```{r}
midGandC %>% 
  filter(q_part %in% c("MidG_ans2", "MidG_ans3")) %>% 
  filter(MidC1_type == "Decreasing and Bounded above") %>% 
  select(study, group, username, q_part, response, MidC1_example) %>% 
  # Use the Ryacas computer algebra system to compare the two answers
  # To prepare, need to add implied multiplication to the expressions
  mutate(
    response_raw = response,
    response = str_replace(response, "2u_n", "2*u_n"),
    response = str_replace(response, "\\)(?=[:digit:])", "\\)\\*"),
    response = str_replace(response, "(?<=[:digit:])\\(", "\\*\\("),
    response = str_replace(response, "(?<=[:digit:])n", "\\*n"),
    response = str_replace(response, "none", "0"),
    response = str_replace(response, "\\[8,6,4,2\\]", "0"),
    response = str_replace(response, "u_\\(n-1\\)", "un"),
    response = str_replace(response, "\\)\\(", "\\)\\*\\("),
    diff = paste0(response, "-(", str_replace(MidC1_example, "(?<=[:digit:])n", "\\*n"), ")")
  ) %>% mutate(
    response_diff = map_chr(diff, ~ .x %>% Ryacas::y_fn("Simplify") %>% Ryacas::yac_str()),
    .keep = "unused"
  ) %>% 
  select(-response) %>% 
  arrange(response_diff) %>% 
  basic_kable()
```

There are 6 cases of the MidC1 example being used verbatim in MidG ans2 or ans3, all by students in classify-first groups.

There are several "near matches" like `5-3n` becoming `2-3n` but it's hard to be definitive in saying that MidC1 was the inspiration for the answer.

#### Session info

```{r}
sessionInfo()
```