---
title: "Study 1 analysis"
author: "George Kinnear"
date: "18/05/2021"
output: 
  html_document:
    code_folding: hide
    df_print: paged
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)

#Plots
library(ggplot2)
library(patchwork) # for combining plots

# For Bayesian stats:
library(bayestestR)
#library(insight)
#library(see)
library(rstanarm)
library(tidybayes)
#library(BayesFactor)
ci_width = 0.95  # Use 95% HDIs
set.seed(20210518)

# Tables
library(knitr)
library(kableExtra)
basic_kable = function(df, ...) {
  df %>% 
    kable(...) %>%
    kable_styling(bootstrap_options = "striped", full_width = F)
}

# Plot styling
heathers = c("#00b3c3","#e51c24","#016637","#fcb123")
heathers = c("#87005a", "#c66005","#016637","#fcb123")
groupcolours = c("C-CG" = heathers[1],
                 "C-GC" = "#6f004a", # darker heathers1
                 "G-CG" = heathers[2],
                 "G-GC" = "#a34e03" # darker heathers2
                 )
four_groups = c("CCG", "CGC", "GCG", "GGC")
theme_set(theme_minimal())
```

For this analysis we use "intention to treat", i.e. we include all students who consented to participate, without regard for whether (or the extent to which) they engaged with the experimental tasks.

```{r read-data, message=FALSE, warning=FALSE}
question_marks = read_csv("data-clean/study1/FAC_LGE-tasks.csv") %>%
  select(question_name, marks_available)

attempt_data_all = read_rds("data-clean/FAC_LGE-attempt_data.rds") %>% 
  # tidy up group names
  mutate(
    group = str_replace(group, "LGE Group ", ""),
    group = str_glue("{str_sub(group, 1, 1)}-{str_sub(group, 2, 3)}"),
    overall_group = str_sub(group, 1, 1)
  ) %>% 
  mutate(
    quiz = case_when(
      question_name %in% c("FarC1", "FarC2") ~ "FT",
      str_detect(question_name, "Mid") ~ "S2",
      TRUE ~ "S1"
    )
  ) %>% 
  left_join(question_marks, by = "question_name") %>% 
  mutate(
    # default to 1 mark per question if not specified
    marks_available = replace_na(marks_available, 1),
    marks_attained = fraction * marks_available
  )
```

# Data summary

```{r}
first_attempts_summary = attempt_data_all %>% 
  filter(attempt_number == 1) %>% 
  group_by(overall_group, group, username, quiz) %>% 
  summarise(
    marks_prop = sum(marks_attained, na.rm = TRUE) / sum(marks_available, na.rm = TRUE),
    n = n(),
    num_nas = sum(is.na(fraction)),
    num_qs = n_distinct(question_name),
    .groups = "drop"
  )

first_attempts_summary %>% 
  group_by(quiz) %>% 
  summarise(
    n_students = n(),
    marks_mean = mean(marks_prop),
    marks_sd = sd(marks_prop),
    .groups = "drop"
  ) %>% 
  basic_kable()

first_attempts_summary %>% 
  group_by(quiz, overall_group, group) %>% 
  summarise(
    n_students = n(),
    marks_mean = mean(marks_prop),
    marks_sd = sd(marks_prop),
    num_qs_in_quiz = max(num_qs),
    overall_nas = mean(num_nas) / num_qs_in_quiz,
    .groups = "drop"
  ) %>% 
  basic_kable()
```

# Item performance

TODO - bring in plots from 3c_check_learning_tasks.R

```{r}
first_attempts = attempt_data_all %>% 
  filter(attempt_number == 1) %>% 
  group_by(quiz, question_name, username) %>% 
  select(quiz, question_name, username, fraction, attempt_number) %>% 
  ungroup()

last_attempts = attempt_data_all %>% 
  group_by(quiz, question_name, username) %>% 
  # select the final attempt
  top_n(n = 1, wt = attempt_number) %>% 
  select(quiz, question_name, username, fraction, attempt_number) %>% 
  ungroup()

best_attempts = attempt_data_all %>% 
  group_by(quiz, question_name, username) %>% 
  # select the best attempt
  top_n(n = 1, wt = fraction) %>% 
  slice(1) %>% 
  select(quiz, question_name, username, fraction, attempt_number) %>% 
  ungroup()

attempts_summary = bind_rows(
  list("first" = first_attempts,
       "last" = last_attempts,
       "best" = best_attempts),
  .id = "attempt_type"
)

attempts_summary_stats = attempts_summary %>% 
  filter(attempt_type != "best") %>% 
  group_by(quiz, question_name, attempt_type) %>% 
  summarise(
    mean_score = mean(fraction, na.rm = TRUE),
    null_response = sum(is.na(fraction)) / n(),
    n = n(),
    .groups = "drop"
  ) %>% 
  ungroup() %>% 
  mutate(
    plot_order = case_when(
      str_detect(question_name, "^C") ~ "1. Learning (C)",
      str_detect(question_name, "^G") ~ "1. Learning (G)",
      str_detect(question_name, "Near") ~ "2. Near",
      str_detect(question_name, "Mid") ~ "3. Mid",
      str_detect(question_name, "Far") ~ "4. Far"
    )
  )

attempts_summary_stats %>% 
  mutate(
    question_name = fct_reorder(question_name, plot_order)
  ) %>% 
  ggplot(aes(x = question_name,
             y = mean_score,
             colour = attempt_type,
             group = attempt_type)) +
  geom_point() +
  geom_line() +
  geom_point(data = attempts_summary_stats,
             mapping = aes(x = question_name,
                 y = null_response),
             shape = 1) +
  facet_wrap(~plot_order, scales = "free_x") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom") +
  ylim(c(0,1)) +
  labs(caption = "Hollow points show the proportion of null responses")
```

```{r}
detailed_scores <- attempt_data_all %>% 
  mutate(
    marks_incdec = case_when(
      !is.na(marks_attained) & str_detect(question_name, "NearC|MidC|FarC") ~ if_else(str_detect(responsesummary, "prt1-1-T"), 1, 0),
      !is.na(marks_attained) & str_detect(question_name, "MidG") ~
        if_else(str_detect(responsesummary, "prt1-2-T"), 1, 0) + if_else(str_detect(responsesummary, "prt2-2-T"), 1, 0)
    ),
    marks_bdd = case_when(
      !is.na(marks_attained) & str_detect(question_name, "NearC|MidC|FarC") ~ if_else(str_detect(responsesummary, "prt2-1-T"), 1, 0),
      !is.na(marks_attained) & str_detect(question_name, "MidG") ~
        if_else(str_detect(responsesummary, "prt3-2-T"), 1, 0) + if_else(str_detect(responsesummary, "prt4-1-T"), 1, 0)
    )

  )

detailed_scores_summary <- detailed_scores %>% 
  filter(attempt_number == 1) %>%
  filter(str_detect(question_name, "Near|Mid|Far")) %>% 
  group_by(userid, group, overall_group) %>% 
  summarise(across(contains("marks"), sum, na.rm = TRUE), .groups = "drop") %>% 
  left_join(
    # add the scores from the Far test only
    detailed_scores %>% 
      filter(attempt_number == 1) %>%
      filter(str_detect(question_name, "Far")) %>% 
      group_by(userid, group, overall_group) %>% 
      summarise(across(c("marks_incdec", "marks_bdd"), sum, na.rm = TRUE), .groups = "drop") %>% 
      rename_with(~ str_replace(., "marks_", "marks_far_")) %>% 
      mutate(marks_far_attained = marks_far_incdec + marks_far_bdd),
    by = c("userid", "group", "overall_group")
  )

detailed_scores_summary %>% head() %>% basic_kable()

detailed_scores_summary %>% 
  pivot_longer(
    cols = contains("marks"),
    names_to = "marks_for",
    values_to = "mark",
    names_prefix = "marks_"
  ) %>% 
  filter(!str_detect(marks_for, "^a")) %>% 
  drop_na() %>% 
  ggplot(aes(x = overall_group, y = mark, colour = marks_for)) +
  geom_violin() +
  ggbeeswarm::geom_beeswarm(alpha = 0.5) +
  facet_wrap(~ marks_for, scales = "free")

detailed_scores_summary %>% 
  mutate(
    first_task = str_sub(group, start = 1, end = 1),
    later_tasks = str_sub(group, start = 3, end = 4)
  ) %>% 
  group_by(overall_group) %>%
  mutate(n = n(), group_label = str_glue("{overall_group} (n={n})")) %>%
  ungroup() %>% 
  write_csv("data-clean/study1_detail.csv")
```


## Initial learning task: Example generation

```{r}
attempts_summary_stats %>% 
  filter(str_detect(question_name, "^G")) %>% 
  ggplot(aes(x = question_name,
             y = mean_score,
             colour = attempt_type,
             group = attempt_type)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1),
        legend.position = "bottom") +
  ylim(c(0,1)) +
  scale_colour_manual("Attempt", values = heathers) +
  labs(y = "Mean proportion of marks attained",
       x = "Question")
```

## Outcome measures

There are various possible outcome measures, based on:

* whether to use all subsequent questions (after the learning task) or only the two questions in the Far test
* whether to focus on the marks relating to increasing/decreasing sequences and boundedness separately

Here we carry out all the analyses to compare the results.

```{r}
compute_contrasts <- function(model) {
  # inspect draws from the model, and compute the contasts of interest
  contrasts <- model %>% 
    tidy_draws() %>% 
    select(starts_with("group"), sigma) %>% 
    mutate(
      CvsG_contrast = (`groupC-CG`+`groupC-GC`)/2 - (`groupG-CG`+`groupG-GC`)/2,
      CvsG_effect = CvsG_contrast / sigma,
      CGvsGC_contrast = (`groupC-CG`+`groupG-CG`)/2 - (`groupC-GC`+`groupG-GC`)/2,
      CGvsGC_effect = CGvsGC_contrast / sigma,
    )
  contrasts_long <- contrasts %>%
    select(-starts_with("group")) %>% 
    pivot_longer(
      cols = contains("vs"),
      names_to = "parameter",
      values_to = "est"
    ) %>% 
    separate(parameter, into = c("comparison", "parameter"))
  
  contrasts_summary <- contrasts_long %>% 
    group_by(comparison, parameter) %>% 
    median_hdi(est) %>% 
    ungroup() %>% 
    mutate(formatted_hdi = str_glue("{round(est, 3)} ({.width*100}% HDI [{round(.lower, 3)}, {round(.upper, 3)}])"))
  
  return(list("contrasts_long" = contrasts_long, "contrasts_summary" = contrasts_summary))
}
```

```{r}
fit_model <- function(data) {
  stan_glm(
    formula = measure ~ 0 + group,
    data = data,
    seed = 123,
    refresh = 0 # to prevent messages about the MCMC iterations being printed
  )
}
all_comparisons <- tibble(
  measure = fct_inorder(c(
    "marks_attained",
    "marks_incdec",
    "marks_bdd",
    "marks_far_attained",
    "marks_far_incdec",
    "marks_far_bdd"
  ))
) %>%
  rowwise() %>%
  mutate(data = map(
    measure,
    ~ detailed_scores_summary %>% select(userid, group, `measure` = all_of(measure))
  )) %>%
  ungroup() %>% 
  mutate(measure_sd = map_dbl(data, ~ .x %>% summarise(marks_sd = sd(measure, na.rm = TRUE)) %>% deframe()))

all_model_fits <- all_comparisons %>% 
  mutate(model = map(data, fit_model))

all_contrasts <- all_model_fits %>% 
  mutate(contrasts = map(model, compute_contrasts)) %>% 
  unnest_wider(contrasts)

all_contrasts %>% 
  select(measure, contrasts_summary) %>% 
  unnest(cols = contrasts_summary) %>% 
  select(measure, comparison, parameter, formatted_hdi) %>% 
  pivot_wider(
    names_from = parameter,
    values_from = formatted_hdi
  ) %>%
  basic_kable()

contrast_data_for_plotting <- all_contrasts %>% 
  select(measure, measure_sd, contrasts_long) %>% 
  unnest(cols = c(contrasts_long)) %>% 
  mutate(rope_width = case_when(parameter == "contrast" ~ 0.1*measure_sd, parameter == "effect" ~ 0.1)) %>% 
  filter(parameter == "effect") %>% 
  mutate(comparison = fct_recode(comparison, "C* vs G*" = "CvsG",  "*CG vs *GC" = "CGvsGC")) %>%
  mutate(far_measure = str_detect(measure, "far")) %>% 
  mutate(
    measure = str_remove_all(measure, "marks_|far_"),
    measure = str_replace_all(
       measure,
       c(
         "attained" = "Marks attained",
         "incdec" = "Marks for\nincreasing/decreasing",
         "bdd" = "Marks for\nboundedness"
       )
     )
  )
  

plot_all <- contrast_data_for_plotting %>%
  filter(!far_measure) %>% 
  ggplot(aes(x = est, y = 0)) +
  geom_rect(aes(xmin = -rope_width, xmax = rope_width), ymin = 0, ymax = 1, fill = "skyblue", alpha = 0.1) +
  geom_vline(xintercept = 0, colour = "grey40") +
  stat_dotsinterval(
    point_interval = mode_hdi,
    .width = .95,
    slab_size = 0,
    quantiles = 100
  ) +
  scale_y_continuous(NULL, breaks = NULL) +
  expand_limits(x = c(-0.1, 0.1)) +
  xlab("Effect size") +
  facet_grid(cols = vars(comparison), rows = vars(measure), scales = "free_x", switch = "y") +
  theme(#panel.grid = element_blank(),
        axis.ticks = element_line(color = "grey40"),
        strip.text.y.left = element_text(angle = 0)
        )

plot_far_only <- contrast_data_for_plotting %>%
  filter(far_measure) %>% 
  ggplot(aes(x = est, y = 0)) +
  geom_rect(aes(xmin = -rope_width, xmax = rope_width), ymin = 0, ymax = 1, fill = "skyblue", alpha = 0.1) +
  geom_vline(xintercept = 0, colour = "grey40") +
  stat_dotsinterval(
    point_interval = mode_hdi,
    .width = .95,
    slab_size = 0,
    quantiles = 100
  ) +
  scale_y_continuous(NULL, breaks = NULL) +
  expand_limits(x = c(-0.1, 0.1)) +
  xlab("Effect size") +
  facet_grid(cols = vars(comparison), rows = vars(measure), scales = "free_x", switch = "y") +
  theme(#panel.grid = element_blank(),
        axis.ticks = element_line(color = "grey40"),
        strip.text.y.left = element_text(angle = 0)
        )

(plot_all + labs(title = "Marks on all questions")) / (plot_far_only + labs(title = "Marks on Far test questions only"))

```

## Main comparison

Since the pattern of results is basically the same for all measures, we focus on `marks_attained` across all questions.

```{r}
model <- (all_model_fits %>% filter(measure == "marks_attained") %>% pull(model))[[1]]
rope_width <- 0.1*(all_model_fits %>% filter(measure == "marks_attained") %>% pull(measure_sd))[[1]]
```

```{r}
means <- modelbased::estimate_means(model, ci = ci_width, centrality = "median", ci_method = "hdi")

detailed_scores_summary %>% 
  ggplot(aes(x = group, y = marks_attained, colour = group)) +
  ggbeeswarm::geom_beeswarm(alpha = 0.5) +
  geom_pointrange(data = means, aes(x = group, y = Mean, ymin = CI_low, ymax = CI_high),
                  position = position_nudge(x = 0.3),
                  size = 0.8) +
  scale_x_discrete("", labels = detailed_scores_summary %>% filter(!is.na(marks_attained)) %>% group_by(group) %>% summarise(group_label = str_glue("{group}\n(n={n()})"), .groups = "drop") %>% deframe(), position = "top") +
  scale_y_continuous("Marks attained (out of 15)") +
  scale_colour_manual("Group", values = groupcolours) +
  theme(legend.position = "none")

ggsave("figs/FIG_study1_marks_attained.pdf", width = 12, height = 7, units = "cm")
```

```{r}
detailed_scores_summary %>% 
  group_by(group) %>% 
  summarise(
    n = sum(!is.na(marks_attained)),
    mean = mean(marks_attained, na.rm = TRUE),
    sd = sd(marks_attained, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  # group is of type "glue" so force it to be a string in order to join on it
  mutate(group = paste(group)) %>% 
  left_join(means, by = "group") %>% 
  # show the HDI but not the estimated mean since that's superfluous
  select(-Mean) %>% 
  mutate(estimate_hdi = str_glue("[{round(CI_low, 3)}, {round(CI_high, 3)}]"), .keep = "unused") %>% 
  basic_kable(caption = "Study 1 group means", digits = 3)
```


```{r compute-contrasts}
contrasts <- model %>% 
  tidy_draws() %>% 
  select(starts_with("group"), sigma) %>% 
  mutate(
    CvsG_contrast = (`groupC-CG`+`groupC-GC`)/2 - (`groupG-CG`+`groupG-GC`)/2,
    CvsG_effect = CvsG_contrast / sigma,
    CGvsGC_contrast = (`groupC-CG`+`groupG-CG`)/2 - (`groupC-GC`+`groupG-GC`)/2,
    CGvsGC_effect = CGvsGC_contrast / sigma,
  )
contrasts_long <- contrasts %>%
  select(-starts_with("group")) %>% 
  pivot_longer(
    cols = contains("vs"),
    names_to = "parameter",
    values_to = "est"
  ) %>% 
  separate(parameter, into = c("contrast", "parameter"))

contrasts_summary <- contrasts_long %>% 
  group_by(contrast, parameter) %>% 
  median_hdi(est) %>% 
  ungroup() %>% 
  mutate(formatted_hdi = str_glue("{round(est, 3)} ({.width*100}% HDI [{round(.lower, 3)}, {round(.upper, 3)}])"))

contrasts_summary %>% 
  select(measure = contrast, parameter, formatted_hdi) %>% 
  pivot_wider(
    names_from = parameter,
    values_from = formatted_hdi
  ) %>%
  basic_kable()

p_diffs <- contrasts_long %>% 
  mutate(key = fct_recode(contrast, "C- vs G-" = "CvsG",  "-CG vs -GC" = "CGvsGC")) %>%
  mutate(key = fct_relevel(key, "C- vs G-")) %>% 
  mutate(rope_width = case_when(parameter == "contrast" ~ rope_width, parameter == "effect" ~ 0.1)) %>% 
  mutate(parameter = fct_recode(parameter, "Raw scores" = "contrast",  "Effect size" = "effect")) %>%
  ggplot(aes(x = est, y = 0)) +
  geom_rect(aes(xmin = -rope_width, xmax = rope_width), ymin = 0, ymax = 1, fill = "skyblue", alpha = 0.1) +
  geom_vline(xintercept = 0, colour = "grey40") +
  stat_dotsinterval(
    point_interval = mode_hdi,
    .width = .95,
    slab_size = 0,
    quantiles = 100
  ) +
  scale_y_continuous(NULL, breaks = NULL) +
  expand_limits(x = c(-0.1, 0.1)) +
  expand_limits(y = c(-0.1)) +
  xlab("Estimated difference") +
  facet_grid(rows = vars(key), cols = vars(parameter), scales = "free_x", switch = "y") +
  theme(#panel.grid = element_blank(),
        axis.ticks = element_line(color = "grey40"),
        strip.text.y.left = element_text(angle = 0)
        )
p_diffs
ggsave("figs/FIG_study1_marks_attained_effect.pdf", width = 14, height = 5, units = "cm")
```


#### Session info

```{r}
sessionInfo()
```