---
title: "Combined analysis of outcome measures for Study 1 and Study 2"
author: "George Kinnear"
date: "09/09/2021"
output: 
  html_document:
    code_folding: hide
    df_print: paged
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)

#Plots
library(ggplot2)
library(patchwork) # for combining plots

# For Bayesian stats:
library(bayestestR)
#library(insight)
#library(see)
library(rstanarm)
library(tidybayes)
#library(BayesFactor)
ci_width = 0.95  # Use 95% HDIs
set.seed(20210518)

# Tables
library(knitr)
library(kableExtra)
basic_kable = function(df, ...) {
  df %>% 
    kable(...) %>%
    kable_styling(bootstrap_options = "striped", full_width = F)
}

# Plot styling
heathers = c("#00b3c3","#e51c24","#016637","#fcb123")
heathers = c("#87005a", "#c66005","#016637","#fcb123")
groupcolours = c("C-CG" = heathers[1],
                 "C-GC" = "#6f004a", # darker heathers1
                 "G-CG" = heathers[2],
                 "G-GC" = "#a34e03" # darker heathers2
                 )
four_groups = c("CCG", "CGC", "GCG", "GGC")
theme_set(theme_minimal())
```

For the comparison of the generate-first and classificy-first groups, the results in Studies 1 and 2 were somewhat different:

* Study 1 found an effect of -0.475 with 95% HDI [−0.804,−0.149]
* Study 2 found an effect of -0.064 with 95% HDI [−0.401, 0.299]

For this analysis, we combine both datasets to produce a single effect size estimate.

```{r read-data, message=FALSE, warning=FALSE}
task_order_data <- bind_rows(
  "study1" = read_csv("data-clean/study1_detail.csv") %>%
              select(username, group = later_tasks, starts_with("marks")),
  "study2" = read_csv("data-clean/study2_detail.csv") %>%
              select(username, group, starts_with("marks")),
  .id = "study"
) %>% 
  mutate(username = paste(study, username, sep = "_"))

task_order_data %>% write_csv("data-clean/studyX_detail.csv")
```

Here we carry out all the analyses to compare the results.

```{r}
compute_contrasts <- function(model) {
  # inspect draws from the model, and compute the contasts of interest
  contrasts <- model %>% 
    tidy_draws() %>% 
    select(starts_with("group"), sigma) %>% 
    mutate(
      CGvsGC_contrast = `groupCG`-`groupGC`,
      CGvsGC_effect = CGvsGC_contrast / sigma,
    )
  contrasts_long <- contrasts %>%
    select(-starts_with("group")) %>% 
    pivot_longer(
      cols = contains("vs"),
      names_to = "parameter",
      values_to = "est"
    ) %>% 
    separate(parameter, into = c("comparison", "parameter"))
  
  contrasts_summary <- contrasts_long %>% 
    group_by(comparison, parameter) %>% 
    median_hdi(est) %>% 
    ungroup() %>% 
    mutate(formatted_hdi = str_glue("{round(est, 3)} ({.width*100}% HDI [{round(.lower, 3)}, {round(.upper, 3)}])"))
  
  return(list("contrasts_long" = contrasts_long, "contrasts_summary" = contrasts_summary))
}
```

```{r}
fit_model <- function(data) {
  stan_glm(
    formula = measure ~ 0 + group,
    data = data,
    seed = 123,
    refresh = 0 # to prevent messages about the MCMC iterations being printed
  )
}
all_comparisons <- tibble(
  measure = fct_inorder(c(
    "marks_attained",
    "marks_incdec",
    "marks_bdd",
    "marks_far_attained",
    "marks_far_incdec",
    "marks_far_bdd"
  ))
) %>%
  rowwise() %>%
  mutate(data = map(
    measure,
    ~ task_order_data %>% select(username, group, `measure` = all_of(measure))
  )) %>%
  ungroup() %>% 
  mutate(measure_sd = map_dbl(data, ~ .x %>% summarise(marks_sd = sd(measure, na.rm = TRUE)) %>% deframe()))

all_model_fits <- all_comparisons %>% 
  mutate(model = map(data, fit_model))

all_contrasts <- all_model_fits %>% 
  mutate(contrasts = map(model, compute_contrasts)) %>% 
  unnest_wider(contrasts)

all_contrasts %>% 
  select(measure, contrasts_summary) %>% 
  unnest(cols = contrasts_summary) %>% 
  select(measure, comparison, parameter, formatted_hdi) %>% 
  pivot_wider(
    names_from = parameter,
    values_from = formatted_hdi
  ) %>%
  basic_kable()

contrast_data_for_plotting <- all_contrasts %>% 
  select(measure, measure_sd, contrasts_long) %>% 
  unnest(cols = c(contrasts_long)) %>% 
  mutate(rope_width = case_when(parameter == "contrast" ~ 0.1*measure_sd, parameter == "effect" ~ 0.1)) %>% 
  filter(parameter == "effect") %>% 
  #mutate(comparison = fct_recode(comparison, "C* vs G*" = "CvsG",  "*CG vs *GC" = "CGvsGC")) %>%
  mutate(far_measure = str_detect(measure, "far")) %>% 
  mutate(
    measure = str_remove_all(measure, "marks_|far_"),
    measure = str_replace_all(
       measure,
       c(
         "attained" = "Marks attained",
         "incdec" = "Marks for\nincreasing/decreasing",
         "bdd" = "Marks for\nboundedness"
       )
     )
  )
  

plot_all <- contrast_data_for_plotting %>%
  filter(!far_measure) %>% 
  ggplot(aes(x = est, y = 0)) +
  geom_rect(aes(xmin = -rope_width, xmax = rope_width), ymin = 0, ymax = 1, fill = "skyblue", alpha = 0.1) +
  geom_vline(xintercept = 0, colour = "grey40") +
  stat_dotsinterval(
    point_interval = mode_hdi,
    .width = .95,
    slab_size = 0,
    quantiles = 100
  ) +
  scale_y_continuous(NULL, breaks = NULL) +
  expand_limits(x = c(-0.1, 0.1)) +
  xlab("Effect size") +
  facet_grid(cols = vars(comparison), rows = vars(measure), scales = "free_x", switch = "y") +
  theme(#panel.grid = element_blank(),
        axis.ticks = element_line(color = "grey40"),
        strip.text.y.left = element_text(angle = 0)
        )

plot_far_only <- contrast_data_for_plotting %>%
  filter(far_measure) %>% 
  ggplot(aes(x = est, y = 0)) +
  geom_rect(aes(xmin = -rope_width, xmax = rope_width), ymin = 0, ymax = 1, fill = "skyblue", alpha = 0.1) +
  geom_vline(xintercept = 0, colour = "grey40") +
  stat_dotsinterval(
    point_interval = mode_hdi,
    .width = .95,
    slab_size = 0,
    quantiles = 100
  ) +
  scale_y_continuous(NULL, breaks = NULL) +
  expand_limits(x = c(-0.1, 0.1)) +
  xlab("Effect size") +
  facet_grid(cols = vars(comparison), rows = vars(measure), scales = "free_x", switch = "y") +
  theme(#panel.grid = element_blank(),
        axis.ticks = element_line(color = "grey40"),
        strip.text.y.left = element_text(angle = 0)
        )

(plot_all + labs(title = "Marks on all questions")) / (plot_far_only + labs(title = "Marks on Far test questions only"))

```

## Main comparison

Since the pattern of results is basically the same for all outcome measures, we focus on `marks_attained` across all questions.

```{r}
model <- (all_model_fits %>% filter(measure == "marks_attained") %>% pull(model))[[1]]
rope_width <- 0.1*(all_model_fits %>% filter(measure == "marks_attained") %>% pull(measure_sd))[[1]]
```

```{r}
means <- modelbased::estimate_means(model, ci = ci_width, centrality = "median", ci_method = "hdi")

task_order_data %>% 
  ggplot(aes(x = group, y = marks_attained)) +
  ggbeeswarm::geom_beeswarm(alpha = 0.5) +
  geom_pointrange(data = means, aes(x = group, y = Mean, ymin = CI_low, ymax = CI_high),
                  position = position_nudge(x = 0.3),
                  size = 0.8) +
  scale_x_discrete("", labels = task_order_data %>% filter(!is.na(marks_attained)) %>% group_by(group) %>% summarise(group_label = str_glue("{group}\n(n={n()})"), .groups = "drop") %>% deframe(), position = "top") +
  scale_y_continuous("Marks attained (out of 15)") +
  theme(legend.position = "none")

ggsave("figs/FIG_studyX_marks_attained.pdf", width = 12, height = 7, units = "cm")
```

```{r}
tab1 <- task_order_data %>% 
  group_by(group) %>% 
  summarise(
    n = sum(!is.na(marks_attained)),
    mean = mean(marks_attained, na.rm = TRUE),
    sd = sd(marks_attained, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  # group is of type "glue" so force it to be a string in order to join on it
  mutate(group = paste(group)) %>% 
  left_join(means, by = "group") %>% 
  # show the HDI but not the estimated mean since that's superfluous
  select(-Mean) %>% 
  mutate(estimate_hdi = str_glue("[{round(CI_low, 3)}, {round(CI_high, 3)}]"), .keep = "unused")

tab1 %>% 
  basic_kable(caption = "Study 1 and 2 combined group means", digits = 3)
```

```{r}
tab1 %>% 
  basic_kable(caption = "Study 1 and 2 combined group means", digits = 3, booktabs = T, format = "latex") %>% 
  cat()
```

```{r}
# inspired by https://stackoverflow.com/questions/61977439/table-below-x-axis-in-ggplot
p1 <- task_order_data %>% 
  ggplot(aes(x = group, y = marks_attained)) +
  ggbeeswarm::geom_beeswarm(alpha = 0.5) +
  geom_pointrange(data = means, aes(x = group, y = Mean, ymin = CI_low, ymax = CI_high),
                  position = position_nudge(x = 0.3),
                  size = 0.8) +
  scale_x_discrete("", labels = task_order_data %>% select(group) %>% distinct(), position = "top") +
  scale_x_discrete("", labels = c("CG" = "Classify first", "GC" = "Generate first"), position = "top") +
  scale_y_continuous("Marks attained (out of 15)") +
  theme(legend.position = "none", axis.text = element_text(size = 11))

p2 <- tab1 %>% 
  mutate(group_label = str_glue("{group}\n(n={n})")) %>% 
  mutate(
    n = str_glue("{n}"),
    mean = str_glue("{round(mean, digits = 3)}"),
    sd = str_glue("{round(sd, digits = 3)}")
  ) %>% 
  pivot_longer(c(n, mean, sd, estimate_hdi), names_to = "layer", values_to = "label") %>% 
  mutate(layer = fct_relevel(layer, c("n", "sd", "mean", "estimate_hdi")) %>% fct_rev()) %>% 
  #filter(layer != "n") %>% 
  ggplot(aes(x = group, y = layer, label = label)) +
  geom_text(size = 10 * 5/14) +
  labs(y = "", x = NULL) +
  scale_y_discrete(labels = c("n" = "N", "mean" = "Mean", "estimate_hdi" = "\n95% HDI\nfor mean")) +
  theme_minimal() +
  theme(axis.line = element_blank(), axis.ticks = element_blank(), axis.text.x = element_blank(),
        panel.grid = element_blank(), strip.text = element_blank())
  
p1 / p2 +  plot_layout(heights = c(8, 2.5))
ggsave("figs/FIG_studyX_marks_attained_with_table.pdf", width = 10, height = 10, units = "cm")
```

```{r compute-contrasts}
contrasts <- model %>% 
  tidy_draws() %>% 
  select(starts_with("group"), sigma) %>% 
  mutate(
    CGvsGC_contrast = `groupCG`-`groupGC`,
    CGvsGC_effect = CGvsGC_contrast / sigma,
  )
contrasts_long <- contrasts %>%
  select(-starts_with("group")) %>% 
  pivot_longer(
    cols = contains("vs"),
    names_to = "parameter",
    values_to = "est"
  ) %>% 
  separate(parameter, into = c("contrast", "parameter"))

contrasts_summary <- contrasts_long %>% 
  group_by(contrast, parameter) %>% 
  median_hdi(est) %>% 
  ungroup() %>% 
  mutate(formatted_hdi = str_glue("{round(est, 3)} ({.width*100}% HDI [{round(.lower, 3)}, {round(.upper, 3)}])"))

contrasts_summary %>% 
  select(measure = contrast, parameter, formatted_hdi) %>% 
  pivot_wider(
    names_from = parameter,
    values_from = formatted_hdi
  ) %>%
  basic_kable()

p_diffs <- contrasts_long %>% 
  mutate(rope_width = case_when(parameter == "contrast" ~ rope_width, parameter == "effect" ~ 0.1)) %>% 
  mutate(parameter = fct_recode(parameter, "Raw scores" = "contrast",  "Effect size" = "effect")) %>%
  ggplot(aes(x = est, y = 0)) +
  geom_rect(aes(xmin = -rope_width, xmax = rope_width), ymin = 0, ymax = 1, fill = "skyblue", alpha = 0.1) +
  geom_vline(xintercept = 0, colour = "grey40") +
  stat_dotsinterval(
    point_interval = mode_hdi,
    .width = .95,
    slab_size = 0,
    quantiles = 100
  ) +
  scale_y_continuous(NULL, breaks = NULL) +
  expand_limits(x = c(-0.1, 0.1)) +
  expand_limits(y = c(-0.1)) +
  xlab("Estimated difference") +
  facet_grid(cols = vars(parameter), scales = "free_x", switch = "y") +
  theme(#panel.grid = element_blank(),
        axis.ticks = element_line(color = "grey40"),
        strip.text.y.left = element_text(angle = 0)
        )
p_diffs
ggsave("figs/FIG_studyX_marks_attained_effect.pdf", width = 14, height = 4, units = "cm")
```


#### Session info

```{r}
sessionInfo()
```